Radial and rotational velocities of young brown dwarfs and very low-mass stars in the Upper Scorpius OB association and the $\rho$~Ophiuchi cloud core

### 2006-05-08 BD_radvel/kurosawa.tex
Revision after Tim's comments.
git-svn-id: https://repository.astro.ex.ac.uk/torus/trunk@713 25464f21-6099-402c-b30a-640f1d40d38b

We present the results of a radial velocity (RV) survey of 14 brown
dwarfs (BDs) and very low-mass (VLM) stars in the Upper Scorpius OB
association (UScoOB) and 3 BD candidates in the $\rho$~Ophiuchi dark
cloud core. We obtained high-resolution echelle spectra at VLT using
Ultraviolet and Visual Echelle Spectrograph (UVES) at two different
epochs for each object, and measured the shifts in their RVs to
identify candidates for binary/multiple systems in the sample. The
average time separation of the RV measurements is 21.6~d, and our
survey is sensitive to the binaries with separation $<0.12$~au.  We
found 2 out of 17 objects (or $24^{+16}_{-14}$~per~cent by fraction)
show a significant RV change in 4--33~d time scale, and are 
considered as binary/multiple `candidates.'  We found no double-lined
spectroscopic binaries in our sample, based on the shape of
cross-corelation curves. The RV dispersion of the objects in UScoOB
is found to be very similar to that of the BD and VLM stars in
Chamaeleon~I (Cha~) previously studies by others. We also found the
distribution of the mean rotational velocities ($v\,\sin i$) of the
UScoOB objects is similar to that of the Cha~I, but the dispersion of
$v\,\sin i$ is significantly larger than that of the Cha~I objects.

@@

\caption{The binary detection probability (solid) as a function of the semimajor
  axis ($a$), based on Monte~Carlo simulations of RV measurements, is
  shown.  For each object in our target list, the simulation was performed
  using the time separation ($\Delta t$) actually used in our
  two-epoch observations (c.f.~Table~\ref{tab:summary_results}) and
  the estimated primary mass  (c.f.~Table~\ref{tab:literatures}).  The
  final detection probability is obtained by averaging
  over the simulated observations of all the objects.  The average
  time separation in the two-epoch observations  is $\sim
  21.6\,\mathrm{d}$, and the average uncertainty in RV is
  $~0.5\,\kmps$. The probability remains fairly
  constant up to $a \approx 0.1\,\mathrm{au}$, and it rapidly decreases
  beyond $a\approx 0.3\,\mathrm{au}$.  The 80~per~cent detection
  probability is achieved for binaries with $a<0.12\,\mathrm{au}$. 
  The binary separation distribution (dash) similar to that of
  \citet{maxted:2005} and \citet{basri:2006} are shown also for a
  comparison. The distribution is a gaussinan in logarithmic
  space, centred at $\log{a}=0.6$, and truncated at $a=15$~au. See text
  for a detail.}

@@

The expected binary fraction for our survey can be obtained
by integrating the product of the detection probability found above
with an assumed binary separation distribution function over a range
of $a$. Here, we simply adopt one of the separation distribution
function used by \citet{maxted:2005}. The binary separation ($a$)
distribution is assumed to be  gaussian in  log($a$) space with its
centre $\log{a_{o}}=0.6$ and width $\sigma_{\log{a}}=1.0$, and
truncated at $\log{a}=1.176$ where $a$ is in au. This form of the
distribution function is motivated by the distribution of solar-type
binaries found in \citet{duquennoy:1991}, and the truncation is
suggested from the lack of wide binaries for $a>15$~au 
(e.g.~\citealt{burgasser:2003}). The distribution is normalized such
that the total binary fraction of 26~per~cent as
suggested by \citet{basri:2006}.  The resulting separation
distribution function is shown also in Fig.~\ref{fig:detect_prob}. 
Using this and the detection probability in the same figure, we find
the expected binary fraction to be $f_{\mathrm{exp}}=3.2$~per~cent
(for $a<0.12$~au), which is
much smaller than our observed binary fraction of
$f_{\mathrm{obs}}=24^{+16}_{-14}$. The two values are not consistent
with each other. There are two possible reasons for the discrepancy:
(1)~the assumed binary separation distribution mentioned above is
unrealistic, and (2)~the uncertainties in our RV measurements are
underestimated.  The first reason is less likely since both
\citet{maxted:2005} and \citet{basri:2006} found this form of the $a$
distribution function is consistent with the observed binary
fractions estimated from many earlier surveys.  We will investigate
the possibility for the second reason in the following.

Apparently, the underestimated uncertainties could have been caused by
neglecting a systematic error. In Section~\ref{sub:Radial-velocities}
(see Fig.~\ref{fig:chisq-test}), we found the histogram of the
$\chi^{2}$ probability $p$, for the RV data fit assuming constant RV,
reasonably matches the expected distribution.  This lead us to
conclude that the uncertainty estimates in RV values were
reasonable. Here, we will inspect the effect of the systematic error
on the observed binary fraction more closely. There are two possible
source of errors: (1)~the RV shifts caused by surface/atmospheric
activity, and (2)~the RV shifts due to the limit of the stability of
autoguiding of the telescope.  \citet{joergens:2006} found that (1) is
negligibly small for a young VLM objects with $M_{*}<0.12\,M_{\sun}$.
The size of RV errors caused by (2) is expected to be $<1\,\kmps$
\citep{bailer-jones:2004}. Note that \citet{maxted:2005} found an
additional systematic error was not needed to be added to the RV
uncertainties of \citet{joergens:2006} who used the same instruments
as ours, and used the same method of estimating the
uncertainties. Hence, we expect the systematic error to be much
smaller than $<1\,\kmps$ if one is required.

The observed binary fraction is expected to be very sensitive to the
assumed size of the systematic error since the RV shifts and the
systematic errors are in same order of magnitude.  This point is
demonstrated further in Fig.~\ref{fig:adding_sys_err} which shows the
the number of observed binaries (non-constant RV objects) plotted
against the size of the additional systematic error.  The number of
detections decreases rapidly as $\sigma_{\mathrm{sys}}$ increases.
The number remains constant at 2 for a relatively wide range of the
systematic errors i.e.  $0.05\,\kmps < \sigma_{\mathrm{sys}} < 0.275\,
\kmps$.  No binary is detected for $\sigma_{\mathrm{sys}} > 
0.375\, \kmps$.  We have computed the $\chi^2$ probabilities for the
RV constant as done earlier (Fig~\ref{fig:chisq-test}) along with the
expected distribution of ($-\log{p}$) with an additional
$\sigma_{\mathrm{sys}}$ introduced in the analysis. We find that the
case with $\sigma_{\mathrm{sys}}=0.1\,\kmps$ gives a better agreement
with the histogram of $-\log{p}$ to the expected
distribution (not shown here), compared to the case with
$\sigma_{\mathrm{sys}}=0\,\kmps$. In the following analysis, we will
adopt $\sigma_{\mathrm{sys}}=0.1\,\kmps$. This means that even if our
systematic error is slightly underestimated, the number of detection
remains the same according to Fig.~\ref{fig:adding_sys_err}.

We have re-analysed the number of observed binaries
(Fig~\ref{fig:chisq-test}), the detection probability
(Fig.~\ref{fig:detect_prob}), and the expected binary fraction as done
earlier but this time with the additional systematic error
$\sigma_{\mathrm{sys}}=0.1\,\kmps$.  The result is summarised in
Table~\ref{tab:binary_fraction}.  
Although the small number statistics makes us difficult to withdraw a
firm conclusion on the binary fraction from our data, we find the case
with $\sigma_{\mathrm{sys}} = 0.1\, \kmps$ has a better agreement with
the expected binary fraction (c.f.~Table~\ref{tab:binary_fraction}).
Our tentative conclusion of the binary fraction found in this analysis
is $f_{\mathrm{sys}}=12^{+13}_{-10}$~per~cent for $a<0.12$~au.  The
large uncertainties in this values is due to small number
statistics. The expected binary fraction
$f_{\mathrm{exp}}=2.9$~per~cent is still within the uncertainty of our
binary fraction.  The two objects identified as 
binary/multiple candidates are GY~310 and USco~104.

\includegraphics[%
  clip,
  scale=0.63]{figures/detect_err_add.eps}

\caption{The number of binaries/multiples (non-constant RV objects)
  detected as a function of the size of additional systematic error ($\sigma_{\mathrm{sys}}$)
  in RV measurements. The number of detections decreases rapidly from 4
  to 2 as $\sigma_{\mathrm{sys}}=0.1\,\kmps$, which may be caused by
  the instability of the autoguiding of the telescope, is added to
  the uncertainties estimated in
  Table~\ref{tab:summary_results}. The number of  detections
  remains constant at 2 for a relatively wide range of the additional
  errors. i.e.~$0.05\,\kmps < \sigma_{\mathrm{sys}} < 0.275\,
  \kmps$. No binary is found for  $\sigma_{\mathrm{sys}} > 0.375\,\kmps$.} 
\label{fig:adding_sys_err}

\begin{table}

\caption{Summary of the observed binary fractions ($f_{\mathrm{obs}}$) compared with the
  expected binary fractions ($f_{\mathrm{exp}}$) with and without a
  systematic error ($\sigma_{\mathrm{sys}}$). }

\label{tab:binary_fraction}

\begin{tabular}{ccccl}
\hline 
$\sigma_{\mathrm{sys}}$&  
N. detected&  
$f_{\mathrm{obs}}$&
$f_{\mathrm{exp}}$&  
obj.~ID  \tabularnewline
($\kmps$)&
&
(per~cent)&
(per~cent)&
\tabularnewline
\hline
$0.0$&
$4$&
$24^{+16}_{-14}$&
$3.2$&
GY~141\tabularnewline
&
&
&
&
GY~310\tabularnewline
&
&
&
&
USco~55\tabularnewline
&
&
&
&
USco~104\tabularnewline
$0.1$&
$2$&
$12^{+13}_{-10}$&
$2.9$&
GY~310\tabularnewline
&
&
&
&
USco~104\tabularnewline
\hline
\end{tabular}

@@

@@

We found 2 (GY~310 and USco~104) out of 17 objects as our final
binary/multiple candidates. This corresponds to the binary fraction of
$12^{+13}_{-10}$~per~cent for the binary separation $a<0.12$~au, which
is larger than the expected binary fraction of 2.9~per~cent, but it is
still within the uncertainties of the observed value.  The expected
binary fraction was computed by using a similar binary separation
models used by \citet{maxted:2005} and \citet{basri:2006}.  We found
that the final estimate of the binary fraction is very sensitive to
the assumed size of systematic error. If the systematic error was
ignored, we would have found 4 binary/multiple candidates.

The recent high-resolution imaging survey of brown dwarfs and very
low-mass objects (M5.5--M7.5 ) in the UScoOB by \citet{kraus:2005} (which was
not known to authors at the time of our observation: April--May, 2004)
confirms that USco~65 and USco~66 are multiple systems, and USco~109
is most likely a multiple system.  Their projected separations are
($a>4.0$~au), and are well above our sensitivity limit of
0.12~au. Hence, it is not surprising that we did not find any of them
as a multiple, and it also indicates that any of them is unlikely to
have a component of a triple system within 0.12~au from their primaries. 
On the other hand, we found USco~104 as a binary candidate, but this
was not in their target list.  Further they found USco~67, 75, 112, 128,
130, 132 as non-multiple, and similarly we found them as
'non-candidates' for multiplicity.

>>>>

We present the results of a radial velocity (RV) survey of 14 brown
dwarfs (BDs) and very low-mass (VLM) stars in the Upper Scorpius OB
association (UScoOB) and 3 BD candidates in the $\rho$~Ophiuchi dark
cloud core. We obtained high-resolution echelle spectra at VLT using
Ultraviolet and Visual Echelle Spectrograph (UVES) at two different
epochs for each object, and measured the shifts in their RVs to
identify candidates for binary/multiple systems in the sample. The
average time separation of the RV measurements is 21.6~d, and our
survey is sensitive to the binaries with separation $<0.12$~au.  We
found 4 out of 17 objects (or $24^{+16}_{-14}$~per~cent by fraction)
show a significant RV change in 4--33~d time scale, and are 
considered as binary/multiple `candidates.'  We found no double-lined
spectroscopic binaries in our sample, based on the shape of
cross-corelation curves. The RV dispersion of the objects in UScoOB
is found to be very similar to that of the BD and VLM stars in
Chamaeleon~I (Cha~) previously studies by others. We also found the
distribution of the mean rotational velocities ($v\,\sin i$) of the
UScoOB objects is similar to that of the Cha~I, but the dispersion of
$v\,\sin i$ is significantly larger than that of the Cha~I objects.

@@

\caption{The binary detection probability (solid) as a function of the semimajor
  axis ($a$), based on Monte~Carlo simulations of RV measurements, is
  shown.  For each object in our target list, the simulation was performed
  using the time separation ($\Delta t$) actually used in our
  two-epoch observations (c.f.~Table~\ref{tab:summary_results}) and
  the estimated primary mass  (c.f.~Table~\ref{tab:literatures}).  The
  final detection probability is obtained by averaging
  over the simulated observations of all the objects.  The average
  time separation in the two-epoch observations  is $\sim
  21.6\,\mathrm{d}$, and the average uncertainty in RV is
  $~0.5\,\kmps$. The probability remains fairly
  constant up to $a \approx 0.1\,\mathrm{au}$, and it rapidly decreases
  beyond $a\approx 0.3\,\mathrm{au}$.  The 80~per~cent detection
  probability is achieved for binaries with $a<0.12\,\mathrm{au}$. }

@@

%% The expected binary fraction for our survey can be obtained
%% by integrating the product of the detection probability found above
%% with an assumed binary separation distribution function over a range
%% of $a$. Here, we simply adopt one of the separation distribution
%% function used by \citet{maxted:2005}. The binary separation ($a$)
%% distribution is assumed to be  gaussian in  log($a$) space with its
%% centre $\log{a_{o}}=0.6$ and width $\sigma_{\log{a}}=1.0$, and
%% truncated at $\log{a}=1.176$ where $a$ is in au. This form of the
%% distribution function is motivated by the distribution of solar-type
%% binaries found in \citet{duquennoy:1991}, and the truncation is
%% suggested from the lack of wide binaries for $a>15$~au 
%% (e.g.~\citealt{burgasser:2003}). The distribution is normalized such
%% that the total binary fraction of 26~per~cent as
%% suggested by \citet{basri:2006}.  The resulting separation
%% distribution function is shown also in Fig.~\ref{fig:detect_prob}. 
%% Using this and the detection probability in the same figure, we find
%% the expected binary fraction to be $f_{\mathrm{exp}}=3.2$~per~cent
%% (for $a<0.12$~au), which is
%% much smaller than our observed binary fraction of
%% $f_{\mathrm{obs}}=24^{+16}_{-14}$. The two values are not consistent
%% with each other. There are two possible reasons for the discrepancy:
%% (1)~the assumed binary separation distribution mentioned above is
%% unrealistic, and (2)~the uncertainties in our RV measurements are
%% underestimated.  The first reason is less likely since both
%% \citet{maxted:2005} and \citet{basri:2006} found this form of the $a$
%% distribution function is consistent with the observed binary
%% fractions estimated from many earlier surveys.  We will investigate
%% the possibility for the second reason in the following.

In Section~\ref{sub:Radial-velocities}
(see Fig.~\ref{fig:chisq-test}), we found the histogram of the
$\chi^{2}$ probability $p$, for the RV data fit assuming constant RV,
reasonably matches the expected distribution.  This lead us to
conclude that the uncertainty estimates in RV values were
reasonable. Here, we will inspect the effect of the systematic error
on the observed binary fraction more closely. There are two possible
source of errors: (1)~the RV shifts caused by surface/atmospheric
activity, and (2)~the RV shifts due to the limit of the stability of
autoguiding of the telescope.  \citet{joergens:2006} found that (1) is
negligibly small for a young VLM objects with $M_{*}<0.12\,M_{\sun}$.
The size of RV errors caused by (2) is expected to be $<1\,\kmps$
\citep{bailer-jones:2004}. Note that \citet{maxted:2005} found an
additional systematic error was not needed to be added to the RV
uncertainties of \citet{joergens:2006} who used the same instruments
as ours, and used the same method of estimating the
uncertainties. Hence, we expect the systematic error to be much
smaller than $<1\,\kmps$ if one is required.

The observed binary fraction is expected to be sensitive to the
assumed size of the systematic error ($\sigma_{\mathrm{sys}}$) since
the RV shifts and the systematic errors are in the same order of
magnitude. The number of binary detections would decrease as
$\sigma_{\mathrm{sys}}$ increases.  We have computed the $\chi^2$
probabilities ($p$) for each object being a RV constant object as done
earlier (Section\ref{sub:Radial-velocities}), but with the assumed 
systematics error of $\sigma_{\mathrm{sys}}=0.2\,\kmps$, which was added to
the uncertainties in the RVs of each object.  The results are
summarised in Table~\ref{tab:prob_compare} along with the probabilities
computed without the systematic error (as in
Section\ref{sub:Radial-velocities}).  The number of objects with
$p<10^{-3}$ decreases from 4 to 2 as the systematic error is
introduced. Unfortunately, the probabilities computed with
$\sigma_{\mathrm{sys}}=0.0\,\kmps$ for GY~141 and USco~55 are just
below $p=10^{-3}$ boundary value; hence, they become RV constant
objects as $\sigma_{\mathrm{sys}}=0.2\,\kmps$ is introduced. This
clearly demonstrates the importance of the uncertainty estimates in
the measured RV values.  We found that the histogram of the $\chi^2$
probabilities with $\sigma_{\mathrm{sys}}=0.2\,\kmps$ (as in
Fig~\ref{fig:chisq-test}), also have a good agreement with the
expected distribution of $-\log{p}$ (not shown here).  
The binary fraction estimated with the assumed systematic error is
$2/17$ or $f=12^{+13}_{-10}$~per~cent for $a<0.12$~au.

%% %==========================================
%% \begin{figure}

%% \begin{center}

%% \includegraphics[%
%%   clip,
%%   scale=0.63]{figures/detect_err_add.eps}

%% \end{center}

%% \caption{The number of binaries/multiples (non-constant RV objects)
%%   detected as a function of the size of additional systematic error ($\sigma_{\mathrm{sys}}$)
%%   in RV measurements. The number of detections decreases rapidly from 4
%%   to 2 as $\sigma_{\mathrm{sys}}=0.1\,\kmps$, which may be caused by
%%   the instability of the autoguiding of the telescope, is added to
%%   the uncertainties estimated in
%%   Table~\ref{tab:summary_results}. The number of  detections
%%   remains constant at 2 for a relatively wide range of the additional
%%   errors. i.e.~$0.05\,\kmps < \sigma_{\mathrm{sys}} < 0.275\,
%%   \kmps$. No binary is found for  $\sigma_{\mathrm{sys}} > 0.375\,\kmps$.} 
%% \label{fig:adding_sys_err}

%% \end{figure}
%% %==========================================

%==========================================
\begin{table}

\caption{Comparison of the $\chi^{2}$ probability ($p$), for each object
  being a RV constant, computed with an addition systematic error
  $\sigma_{\mathrm{sys}}$ (right), and without (left). The objects are
  listed in ascending order of $p$. The ellipses represent the
  boundary ($p=10^{-3}$) between RV constant objects and RV 
  variable objects. }

\label{tab:prob_compare}

\begin{center}
\begin{tabular}{lclc}
\hline 
&
$\sigma_{\mathrm{sys}}=0.0\,\kmps$&
&
$\sigma_{\mathrm{sys}}=0.2\,\kmps$\tabularnewline
obj.~ID&
$p$&
obj.~ID&
$p$\tabularnewline
\hline
USco~104&
$4.3\times10^{-12}$&
USco~104&
$3.5\times10^{-6}$\tabularnewline
GY~310&
$1.8\times10^{-10}$&
GY~310&
$2.0\times10^{-5}$\tabularnewline
GY~141&
$1.6\times10^{-4}$&
$\cdots$&
$\cdots$\tabularnewline
USco~55&
$7.2\times10^{-4}$&
GY~141&
$5.8\times10^{-2}$\tabularnewline
$\cdots$&
$\cdots$&
USco~55&
$1.6\times10^{-1}$\tabularnewline
USco~101&
$8.9\times10^{-2}$&
USco~130&
$2.3\times10^{-1}$\tabularnewline
USco~66&
$9.2\times10^{-2}$&
USco~132&
$2.6\times10^{-1}$\tabularnewline
USco~75&
$2.8\times10^{-1}$&
USco~101 &
$2.8\times10^{-1}$\tabularnewline
USco~109&
$3.0\times10^{-1}$&
GY~5&
$2.8\times10^{-1}$\tabularnewline
USco~112&
$3.1\times10^{-1}$&
USco~75&
$3.9\times10^{-1}$\tabularnewline
USco~132&
$3.2\times10^{-1}$&
USco~66&
$4.3\times10^{-1}$\tabularnewline
USco-~30&
$3.2\times10^{-1}$&
USco~128&
$4.6\times10^{-1}$\tabularnewline
GY~5&
$3.7\times10^{-1}$&
USco~53&
$5.8\times10^{-1}$\tabularnewline
USco~53 &
$4.2\times10^{-1}$&
USco40&
$6.0\times10^{-1}$\tabularnewline
USco~67&
$4.8\times10^{-1}$&
USco~112&
$6.5\times10^{-1}$\tabularnewline
USco~128&
$5.4\times10^{-1}$&
USco~100&
$6.6\times10^{-1}$\tabularnewline
USco~100&
$6.2\times10^{-1}$&
USco~67&
$7.0\times10^{-1}$\tabularnewline
USco~40&
$7.2\times10^{-1}$&
USco~109&
$7.1\times10^{-1}$\tabularnewline
\hline
\end{tabular}

@@

%% %==========================================
%% \begin{table}

%% \caption{Summary of the observed binary fractions ($f_{\mathrm{obs}}$) compared with the
%%   expected binary fractions ($f_{\mathrm{exp}}$) with and without a
%%   systematic error ($\sigma_{\mathrm{sys}}$). }

%% \label{tab:binary_fraction}

%% \begin{center}

%% \begin{tabular}{ccccl}
%% \hline 
%% $\sigma_{\mathrm{sys}}$&  
%% N. detected&  
%% $f_{\mathrm{obs}}$&
%% $f_{\mathrm{exp}}$&  
%% obj.~ID  \tabularnewline
%% ($\kmps$)&
%% &
%% (per~cent)&
%% (per~cent)&
%% \tabularnewline
%% \hline
%% $0.0$&
%% $4$&
%% $24^{+16}_{-14}$&
%% $3.2$&
%% GY~141\tabularnewline
%% &
%% &
%% &
%% &
%% GY~310\tabularnewline
%% &
%% &
%% &
%% &
%% USco~55\tabularnewline
%% &
%% &
%% &
%% &
%% USco~104\tabularnewline
%% $0.1$&
%% $2$&
%% $12^{+13}_{-10}$&
%% $2.9$&
%% GY~310\tabularnewline
%% &
%% &
%% &
%% &
%% USco~104\tabularnewline
%% \hline
%% \end{tabular}

%% \end{center}

%% \end{table}
%==========================================

@@

We found 4 (GY~141, GY~310, USco~55 and USco~104) out of 17 objects as our 
binary/multiple candidates. This corresponds to the binary fraction of
$24^{+16}_{-14}$~per~cent for the binary separation $a<0.12$~au.  We found
that the estimated binary fraction is very sensitive to
the assumed size of systematic error. If an additional systematic
error of $0.2\,\kmps$  was added to the uncertainties in the RV
measurements, we found only 2 (GY~310 and USco~104) out of 17 objects
as our binary/multiple candidates, which corresponds to the binary
fraction of $12^{+13}_{-10}$~per~cent.

The recent high-resolution imaging survey of brown dwarfs and very
low-mass objects (M5.5--M7.5 ) in the UScoOB by \citet{kraus:2005} (which was
not known to authors at the time of our observation: April--May, 2004)
confirms that USco~55 and USco~66 are multiple systems, and USco~109
is most likely a multiple system.  Their projected separations are
($a>4.0$~au), and are well above our sensitivity limit of
0.12~au. Hence, it is not surprising that we did not find USco~66 and
USco~109 as multiples. Interestingly, we also found USco~55 as a
candidate for a multiple system. This may indicate that USco~55 is  
a triple system with one of object located within 0.12~au from the
primary.  In addition, we found USco~104 as a binary candidate, but this
was not in their target list.  Further they found USco~67, 75, 112, 128,
130, 132 as non-multiple, and similarly we found them as
'non-candidates' for multiplicity.
